Chemical tagging can work: \\ Identification of stellar phase-space structures \\ purely by chemical-abundance similarity\vspace{-7ex}

### 2016-08-07 papers/kmeans/kmeans.tex
Removed wrong paragraph on Pal5 and associated references. Added better paragraph for M5

We chose a few interesting cases from the high-density clusters
in the $K=256$ run and show them in
\figurename~\ref{fig:M13}, \ref{fig:Pal5}, \ref{fig:Sgr},
\ref{fig:halo}, and \ref{fig:disk}.
The first three of these are dominated by stars in the known clusters
M13 and M5, and the Sagittarius stream (we identify overlap with these objects
by looking at stellar position and velocity, and, in some cases,
\apogee\ targeting flags).
The fourth is a halo structure with high velocity dispersion and
possibly accreted.
The fifth is a thin-disk star-formation feature.
We will discuss the astrophysical implications of these results in
\sectionname~\ref{sec:discussion}.

@@

Although we show only three stellar clusters in
\figurename~\ref{fig:M13}, \ref{fig:Pal5}, and \ref{fig:Sgr}, many
other clusters are visible among our densest abundance-space k-means
clusters.
These include M15, M92, and M107, among others.
We haven't asked yet whether we detect---as abundance-space
overdensities---all of the stellar clusters we expect to find. (Many
clusters within the \apogee\ data set only contain a few plausible 
members, thereby  complicating any inferences we wish to make about 
detection completeness using simple k-means.) 
This \documentname\ is simply a demonstration of the chemical-tagging
concept; a full investigation of whether we can construct complete
catalogs of stellar clusters is beyond our present scope.

\begin{figure}[!bp]
\includegraphics[width=1.5\figwidth]{./figs/clusters_0256.png}
\caption{The distribution of membership and density for the 256
  abundance-space clusters returned by the k-means algorithm at
  $K=256$.  The densest cluster (which is displayed in more detail in 
  \figurename~\ref{fig:M13}) is far more dense than the second
  densest.\label{fig:densities}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0253}
\caption{Same as \figurename~\ref{fig:all} except that the entire
  sample has been made gray and the members of the most dense $K=256$
  abundance-space cluster have been rendered as unique, prominent
  triangles (color is velocity rank, orientation is $\logg$ rank).
  The lower-right plot shows this cluster (circle) in context of the
  other clusters (dots).  This cluster, which was identified only in
  abundance space (six projections of which are the top six panels in
  this \figurename), turns out to be dominated by the halo globular
  cluster M13. The symbol orientations and colors have no meaning;
  they are randomly generated---one unique color and orientation for
  each highlighted star---to make the points cross-identifiable across
  panels.\label{fig:M13}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0034}
\caption{Same as \figurename~\ref{fig:M13} but for another dense
  abundance-space cluster.
  This one turns out to be dominated by globular
  cluster M5.\label{fig:Pal5}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0177}
\caption{Same as \figurename~\ref{fig:M13} but for another dense
  abundance-space cluster.
  This one turns out to be dominated by the Sagittarius dwarf spheroidal galaxy.\label{fig:Sgr}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0010}
\caption{Same as \figurename~\ref{fig:M13} but for another dense
  abundance-space cluster.
  This one turns out to be dominated by a hitherto unrecognized high-velocity-dispersion
  structure in the Galaxy halo.\label{fig:halo}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0141}
\caption{Same as \figurename~\ref{fig:M13} but for another dense
  abundance-space cluster.
  This one turns out to be dominated by a very thin young stellar
  structure in the Galaxy disk, probably one that has been discovered previously \citep{wegg}.\label{fig:disk}}
\end{figure}

@@

% HOGG: THIS PARAGRAPH IS ALL WRONG!
% There are scant abundance measurements reported for Palomar 5, for which
% most stars in the cluster in \figurename~\ref{fig:Pal5} belong. We find a mean
% metallicity of [Fe/H]$ = -1.3$, which lies between the only
% photometric \citep{Geisler_1997} and spectroscopic \citep{Smith_1985,Smith_2002} abundance
% measurements for this cluster.  The anti-correlation in C--N and Na--O abundance
% have been seen elsewhere \citep[][respectively]{Smith_1985,Smith_2002}, although
% to our knowledge the Mg--Al anti-correlation has not previously been shown for this cluster.
% Indeed, this work contains the largest number of abundances for Palomar 5,
% permitting a thorough search of tidally disrupted members \citep[for example,][]{Kuzma_2015} 
% that have retained the unique cluster abundance signature.

[HOGG, CASEY: INSERT PARAGRAPH ABOUT M5 HERE]

@@

It is worth reiterating here what is written above:
There is no sense at all in which \figurename~\ref{fig:M13},
\ref{fig:Pal5}, \ref{fig:Sgr}, \ref{fig:halo}, and \ref{fig:disk} show
a representative or complete set of features found in abundance space.
These features were hand-chosen to be obviously interesting and
interpetable.
There are many other things to be found in this data set, and many
more effective models to be built of abundance-space structure.
The only strong conclusion of this investigation is that chemical
tagging \emph{can} work; the abundance measurements with \thecannon\ are
precise enough and the chemical-abundance space is informative enough.

@@

\bibitem[Geisler \etal(1997)]{Geisler_1997}
Geisler, D., Claria, J.~J., Minniti, D., 1997, \pasp, 109, 799 

@@

@@

@@

\bibitem[Kuzma \etal(2015)]{Kuzma_2015}
Kuzma, P.~B., Da Costa, G.~S., Keller, S.~C., Maunder, E., 2015, \mnras, 446, 3297 

@@

\bibitem[Smith(1985)]{Smith_1985}
Smith, G.~H., 1985, \apj, 298, 249 
\bibitem[Smith \etal(2002)]{Smith_2002}
Smith, G.~H., Sneden, C., Kraft, R.~P., 2002, \aj, 123, 1502 

>>>>

We chose a few interesting cases from the high-density clusters
in the $K=256$ run and show them in
\figurename~\ref{fig:M13}, \ref{fig:M5}, \ref{fig:Sgr},
\ref{fig:halo}, and \ref{fig:disk}.
The first three of these are dominated by stars in the known clusters
M13 and M5, and the Sagittarius stream (we identify overlap with these objects
by looking at stellar position and velocity, and, in some cases,
\apogee\ targeting flags).
The fourth is a halo structure with high velocity dispersion and
possibly accreted.
The fifth is a thin-disk star-formation feature.
We will discuss the astrophysical implications of these results in
\sectionname~\ref{sec:discussion}.

@@

Although we show only three stellar clusters in
\figurename~\ref{fig:M13}, \ref{fig:M5}, and \ref{fig:Sgr}, many
other clusters are visible among our densest abundance-space k-means
clusters.
These include M15, M92, and M107, among others.
We haven't asked yet whether we detect---as abundance-space
overdensities---all of the stellar clusters we expect to find. (Many
clusters within the \apogee\ data set only contain a few plausible 
members, thereby  complicating any inferences we wish to make about 
detection completeness using simple k-means.) 
This \documentname\ is simply a demonstration of the chemical-tagging
concept; a full investigation of whether we can construct complete
catalogs of stellar clusters is beyond our present scope.

\begin{figure}[!bp]
\includegraphics[width=1.5\figwidth]{./figs/clusters_0256.png}
\caption{The distribution of membership and density for the 256
  abundance-space clusters returned by the k-means algorithm at
  $K=256$.  The densest cluster (which is displayed in more detail in 
  \figurename~\ref{fig:M13}) is far more dense than the second
  densest.\label{fig:densities}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0253}
\caption{Same as \figurename~\ref{fig:all} except that the entire
  sample has been made gray and the members of the most dense $K=256$
  abundance-space cluster have been rendered as unique, prominent
  triangles (color is velocity rank, orientation is $\logg$ rank).
  The lower-right plot shows this cluster (circle) in context of the
  other clusters (dots).  This cluster, which was identified only in
  abundance space (six projections of which are the top six panels in
  this \figurename), turns out to be dominated by the halo globular
  cluster M13. The symbol orientations and colors have no meaning;
  they are randomly generated---one unique color and orientation for
  each highlighted star---to make the points cross-identifiable across
  panels.\label{fig:M13}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0034}
\caption{Same as \figurename~\ref{fig:M13} but for another dense
  abundance-space cluster.
  This one turns out to be dominated by globular
  cluster M5.\label{fig:M5}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0177}
\caption{Same as \figurename~\ref{fig:M13} but for another dense
  abundance-space cluster.
  This one turns out to be dominated by the Sagittarius dwarf spheroidal galaxy.\label{fig:Sgr}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0010}
\caption{Same as \figurename~\ref{fig:M13} but for another dense
  abundance-space cluster.
  This one turns out to be dominated by a hitherto unrecognized high-velocity-dispersion
  structure in the Galaxy halo.\label{fig:halo}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0141}
\caption{Same as \figurename~\ref{fig:M13} but for another dense
  abundance-space cluster.
  This one turns out to be dominated by a very thin young stellar
  structure in the Galaxy disk, probably one that has been discovered previously \citep{wegg}.\label{fig:disk}}
\end{figure}

@@

We find a mean metallicity of [Fe/H]$ = -1.3$ for the cluster we associate as M~5 (\figurename~\ref{fig:M5}), which agrees well with the [Fe/H]$ = -1.33 \pm 0.03$ measurement reported by \citet{Koch_2010}, and the $-1.29 \pm 0.02$ value listed in the compilation of \citet[][accessed 2016]{Harris_1996}.  The correlations in light elements---specifically Mg--Al and C--N---also agree well with other studies \citep{Ivans_2001,meszaros}.  In particular we find only a weak correlation in the [Na/Fe]--[O/Fe] abundance ratios \citep{Lai_2011}.  However, when we consider the extent of the literature on M~5, it would suggest that we do not recover the full extent of these correlations: the stars with the highest [Na/Fe], [Mg/Al], and lowest [O/Fe] abundance ratios are not represented in the cluster that we associate as M~5.  This is likely a consequence of our (poor) choice of the k-means algorithm, which is most effective for near-normal distributions, and clearly less powerful for clusters (in a statistical sense) that have polynomial relationships in dimensional space (e.g., [Na/Fe]--[O/Fe]).  Indeed, \citet{meszaros} report abundances for 122 members of M~5 using \apogee\ \acronym{DR12} data, whereas the cluster we associate as M~5 only contains $\approx60$ members, and only covers about half of the extent of [Na/Fe]--[O/Fe] relationship.

@@

It is worth reiterating here what is written above:
There is no sense at all in which \figurename~\ref{fig:M13},
\ref{fig:M5}, \ref{fig:Sgr}, \ref{fig:halo}, and \ref{fig:disk} show
a representative or complete set of features found in abundance space.
These features were hand-chosen to be obviously interesting and
interpetable.
There are many other things to be found in this data set, and many
more effective models to be built of abundance-space structure.
The only strong conclusion of this investigation is that chemical
tagging \emph{can} work; the abundance measurements with \thecannon\ are
precise enough and the chemical-abundance space is informative enough.

@@

@@

\bibitem[Ivans et al.(2001)]{Ivans_2001} Ivans, I.~I., Kraft, R.~P., Sneden, C., et al.\ 2001, \aj, 122, 1438 

@@

\bibitem[Koch \& McWilliam(2010)]{Koch_2010} Koch, A., \& McWilliam, A.\ 2010, \aj, 139, 2289 

@@

\bibitem[Lai et al.(2011)]{Lai_2011} Lai, D.~K., Smith, G.~H., Bolte, M., et al.\ 2011, \aj, 141, 62 

@@
